---
title: "deniedthevoteFIX"
author: "Joseph Pacifico"
date: "1/6/2021"
output: html_document
---

setwd("D:/Libraries 1/Documents/QMSS/Data For Good/Tableau")

```{r}
library(tidyverse)
library(xlsx)
#library(xlsx)
#library(openxlsx)

setwd("D:/Libraries 1/Documents/QMSS/Data For Good/Tableau")
denied <- openxlsx::read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\1870 denied the vote data.xlsx", sheet="usa_00003", startRow = 1,rows= 1:125990, cols = 1:32)

# to match codebook codes format
denied$CITY <- sprintf("%04d", as.numeric(denied$CITY)) # 4 zeros
denied$STATEICP <- sprintf("%02d", as.numeric(denied$STATEICP)) # 2

denied$RACED <- sprintf("%03d", as.numeric(denied$RACED)) # 3
denied$RACED <- as.character(denied$RACED)

denied$BPLD <-  sprintf("%05d", as.numeric(denied$BPLD)) # 5 zeros
denied$BPLD <- as.character(denied$BPLD)
# to match codebook codes format

denied$REGION <- as.character(denied$REGION)
denied$CITY <- as.character(denied$CITY)
denied$STATEICP <- as.character(denied$STATEICP)
denied$RACESINGD <- as.character(denied$RACESINGD)


denied

```




# regionFIXED 2 zeros
```{r}

region <- read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'division', rowIndex= 1:13, colIndex = 1:2) 
region <- region %>% plyr::rename(c('REGION.Census.region.and.division' = 'region', 'NA.' = 'regionlocation'))

region <- as.data.frame(str_split_fixed(region$region, " ", 2)) %>% plyr::rename(c('V1' = 'regioncode', 'V2' = 'regionname')) %>% mutate(regionlocation = region$regionlocation)
#region$regioncode <- as.double(region$regioncode)
region

```



# cityFIXED 4 zeros
```{r}

city <- read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'city', rowIndex= 1:1172, colIndex = 1:2, colClasses='character')
city <- city %>% plyr::rename(c('code' = 'citycode', 'cita' = 'citystate'))

city$citycode <- sprintf("%04d", as.numeric(city$citycode))
#city$citycode <- as.character(city$citycode)
#city$cityboth <- paste(city$citycode, city$citystate, sep = " ")
city


#https://stackoverflow.com/questions/5812493/how-to-add-leading-zeros
```







1. https://stackoverflow.com/questions/61256777/replace-region-numbers-with-names
2. https://stackoverflow.com/questions/7547597/dictionary-style-replace-multiple-items

https://stackoverflow.com/questions/50224076/pass-variables-from-another-dataframe-in-a-case-when-statement-with-dplyr-in-r


https://stackoverflow.com/questions/41671856/error-incompatible-size-when-mutating-in-dplyr

# stateicpFIXED 2 zeros
```{r}

state <- read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'state', rowIndex= 1:56, colIndex = 1:2)
state <- state %>% plyr::rename(c('STATEICP.State..ICPSR.code.' = 'stateicp'))
state <- as.data.frame(str_split_fixed(state$stateicp, " ", 2)) %>% plyr::rename(c('V1' = 'statecode', 'V2' = 'statename'))

state$statecode <- sprintf("%02d", as.numeric(state$statecode)) #adds zeroes to match the same format as in the codebook

state

#https://stackoverflow.com/questions/44488300/change-column-variable-string-from-1-to-01
```


# TEST (CAN DISREGARD) 
```{r}



test <- select(denied, YEAR, SAMPLE, SERIAL, CITY, REGION, STATEICP)
test$CITY <- as.character(test$CITY)
test$REGION <- as.character(test$REGION)
test$STATEICP <- as.character(test$STATEICP)
test

```





# racedFIXED 3 zeros
```{r}

raced <- read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'raced', rowIndex= 1:117, colIndex = 1)
raced <- raced %>% plyr::rename(c('RACED.Race..detailed.version.' = 'raced'))


raced <- as.data.frame(str_split_fixed(raced$raced, " ", 2)) %>% plyr::rename(c('V1' = 'racedcode', 'V2' = 'racedname'))

#city$citycode <- as.character(city$citycode)
#city$cityboth <- paste(city$citycode, city$citystate, sep = " ")



raced2 <- read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'raced2', rowIndex= 1:131, colIndex = 1:2)
raced2 <- raced2 %>% plyr::rename(c('raced2' = 'racedcode', 'NA.' = 'racedname'))

raced2$racedcode <- as.character(raced2$racedcode)

raced_final <- full_join(raced, raced2) %>% arrange(racedcode)

raced_final <- na.omit(raced_final)
raced_final
```


# bpldFIXED 5 zeros
```{r}

bpld <- read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'bpld', rowIndex= 1:38, colIndex = 1)
bpld <- bpld %>% plyr::rename(c('BPLD.Birthplace..detailed.version...UNITED.STATES' = 'bpld'))


bpld <- as.data.frame(str_split_fixed(bpld$bpld, " ", 2)) %>% plyr::rename(c('V1' = 'bpldcode', 'V2' = 'bpldname'))

#city$citycode <- as.character(city$citycode)
#city$cityboth <- paste(city$citycode, city$citystate, sep = " ")



bpld2 <- xlsx::read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'bpld2', rowIndex= 1:535, colIndex = 1:2, colClasses='character')
bpld2 <- bpld2 %>% plyr::rename(c('title' = 'bpldcode', 'NA.' = 'bpldname'))




bpld2$bpldcode <- sprintf("%05d", as.numeric(bpld2$bpldcode)) #adds zeroes to match the same format as in the codebook


bpld2$bpldcode <- as.character(bpld2$bpldcode)

bpld_final <- full_join(bpld, bpld2) 

bpld_final <- arrange(bpld_final, bpldcode) %>% slice(-c(1:33)) 
bpld_final
```
# racesingdFIXED 
```{r}

racesingd <- read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'racesingd', rowIndex= 1:18, colIndex = 1)

racesingd <- racesingd %>% plyr::rename(c('RACESINGD.Race..Single.race.identification..detailed.version.' = 'racesingd'))


racesingd <- as.data.frame(str_split_fixed(racesingd$racesingd, " ", 2)) %>% plyr::rename(c('V1' = 'racesingdcode', 'V2' = 'racesingdname'))

racesingd
```



# citizen 
```{r}

citizen <- read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\Tableau\\deniedthevoteCODEBOOKFIX.xlsx", sheetName = 'citizen', rowIndex= 1:8, colIndex = 1)

citizen <- citizen %>% plyr::rename(c('CITIZEN.Citizenship.status' = 'citizen'))


citizen <- as.data.frame(str_split_fixed(citizen$citizen, " ", 2)) %>% plyr::rename(c('V1' = 'citizencode', 'V2' = 'citizenstatus'))

citizen
```


# TESTING VARIABLE MERGE (CAN DISREGARD) add in BPLD, RACED and RACESINGD CITIZEN to deniedEXPORT
```{r}

#RACED, RACESINGD, BPLD

test <- select(denied, YEAR, SAMPLE, SERIAL, RACED, RACESINGD, BPLD, CITIZEN)


test$RACED <- sprintf("%03d", as.numeric(test$RACED)) #adds zeroes to match the same format as in the codebook
test$RACED <- as.character(test$RACED)

test$BPLD <-  sprintf("%05d", as.numeric(test$BPLD)) #adds zeroes to match the same format as in the codebook
test$BPLD <- as.character(test$BPLD)

test$RACESINGD <- as.character(test$RACESINGD)

test$CITIZEN <- as.character(test$CITIZEN)

test <- left_join(test, raced_final, by = c("RACED" = "racedcode")) %>% 
  left_join(racesingd, by = c("RACESINGD" = "racesingdcode")) %>% 
  left_join(bpld_final, by = c("BPLD" = "bpldcode")) %>% 
  left_join(citizen, by = c("CITIZEN" = "citizencode"))
test
```



https://stackoverflow.com/questions/18279268/read-xlsx-and-colclasses



```{r}
denied$REGION <- as.character(denied$REGION)
denied$CITY <- as.character(denied$CITY)
denied$STATEICP <- as.character(denied$STATEICP)
denied$CITIZEN <- as.character(denied$CITIZEN)
#https://stackoverflow.com/questions/61256777/replace-region-numbers-with-names
```


# add in BPLD, RACED and RACESINGD CITIZEN to deniedEXPORT
```{r}


deniedFIXEDexport <- left_join(denied, region, by = c("REGION" = "regioncode")) %>% 
  left_join(city, by = c("CITY" = "citycode")) %>% 
  left_join(state, by = c("STATEICP" = "statecode")) %>% 
  left_join(raced_final, by = c("RACED" = "racedcode")) %>% 
  left_join(racesingd, by = c("RACESINGD" = "racesingdcode")) %>% 
  left_join(bpld_final, by = c("BPLD" = "bpldcode")) %>% 
  left_join(citizen, by = c("CITIZEN" = "citizencode"))

deniedFIXEDexport
```




# total re-coded variables
regionname
regionlocation
citystate
statename
racedname
racesingdname
bpldname
citizen

```{r}

select(deniedFIXEDexport, 
       REGION,
       regionname,
       regionlocation,
       CITY,
       citystate,
       STATEICP,
       statename, 
       BPLD,
       bpldname,
       RACED,
       racedname,
       RACESINGD, 
       racesingdname, 
       CITIZEN, 
       citizenstatus
       )


#https://stackoverflow.com/questions/61256777/replace-region-numbers-with-names
```


```{r}

unique(deniedFIXEDexport$citizenstatus)

```







#export the data to tableau
```{r}

#export to csv
#write.csv(x=deniedFIXEDexport, file="1870 denied the vote dataFIXED.csv")


```



```{r}
df4 <- read.csv("1870 denied the vote dataFIXED.csv")
df4
```





 I assume even Excel could do the basic statistics of the total and by state — race, birthplace, age averages, to figure out what we have here.

```{r}


df5 <- df4 %>% 
  select(SERIAL, REGION, STATEICP, CITY, COUNTYNHG, SEX, AGE, RACE, RACED, BPL, BPLD, CITIZEN, RACESING, RACESINGD, VOTE, regionname, regionlocation, citystate, statename, racedname, racesingdname, bpldname, citizenstatus )


df5
```


# Age by State 
```{r}
age_state <- df5 %>% 
  group_by(statename, SEX) %>%
  summarize(avg_age = mean(AGE, na.rm=TRUE)) %>% 
  mutate(Gender = case_when(SEX == 1 ~ "Male"))
age_state

#write.csv(age_state, file="age_state.csv")
```

- only males in the sample. 



# Age by City
```{r}
age_city <- df5 %>% 
  group_by(citystate, SEX) %>%
  summarize(avg_age = mean(AGE, na.rm=TRUE)) %>% 
  mutate(Gender = case_when(SEX == 1 ~ "Male"))

age_city

#write.csv(age_city, file="age_city.csv")
```

# how many in named cities vs not 
```{r}

named_city_totals <- df5 %>% 
  group_by(citystate) %>%
  summarize(total = n()) 

named_city_totals
#write.csv(named_city_totals, file="named_city_totals.csv")
```

# how many in named cities vs not 
```{r}

named_vs_not_city <- df5 %>% 
  group_by(citystate) %>%
  summarize(total = n()) %>%
  mutate(named = case_when(
    citystate == 'Not in identifiable city (or size group)' ~ 'Not in identifiable city (or size group)',
    citystate != 'Not in identifiable city (or size group)'~ 'named city'
  )) %>% 
  group_by(named) %>% 
  summarize(total_named = sum(total))

named_vs_not_city
#write.csv(named_vs_not_city, file="named_vs_not_city.csv")
```

# how many in named states vs not 
```{r}

named_state_totals <- df5 %>% 
  group_by(statename) %>%
  summarize(total = n()) 

named_state_totals
#write.csv(named_state_totals, file="named_state_totals.csv")
```

# how many in named states vs not 
```{r}

named_vs_not_state <- df5 %>% 
  group_by(statename) %>%
  summarize(total = n()) %>%
  summarize(totals = sum(total))

named_vs_not_state
#write.csv(named_vs_not_state, file="named_vs_not_state.csv")
```

# Race by State
```{r}
df5$racedname <- factor(df5$racedname)

race_state <- df5 %>% 
  group_by(statename, racedname) %>%
  summarize(total = n()) 

race_state
#write.csv(race_state, file="race_state.csv")
```

# Race by City
```{r}

race_city <- df5 %>% 
  group_by(citystate, racedname) %>%
  summarize(total = n()) 

race_city
#write.csv(race_city, file="race_city.csv")
```



# birthplace by city 
```{r}

birthplace_city <- df5 %>% 
  group_by(citystate, bpldname) %>%
  summarize(total = n()) 

birthplace_city
#write.csv(birthplace_city, file="birthplace_city.csv")
```

# birthplace by State 
```{r}

birthplace_state <- df5 %>% 
  group_by(statename, bpldname) %>%
  summarize(total = n()) 

birthplace_state
#write.csv(birthplace_state, file="birthplace_state.csv")
```


# race_totals
```{r}

race_totals <- df5 %>% 
  group_by(RACED, racedname) %>%
  summarize(total = n()) 

race_totals
#write.csv(race_totals, file="race_totals.csv")
```


# age_totals
```{r}
df5$SERIAL <- as.character(df5$SERIAL)



age_totals <- df5 %>% 
  select(SERIAL, AGE) %>% 
  mutate(age_range = case_when(
    AGE >= 21 & AGE <= 30 ~ "age21_30", 
    AGE >= 31 & AGE <= 40 ~ "age31_40",
    AGE >= 41 & AGE <= 50 ~ "age41_50",
    AGE >= 51 & AGE <= 60 ~ "age51_60",
    AGE >= 61 & AGE <= 70 ~ "age61_70",
    AGE >= 71 & AGE <= 80 ~ "age71_80",
    AGE >= 81 & AGE <= 90 ~ "age81_90",
    AGE >= 91 & AGE <= 100 ~ "age91_100", 
    AGE >= 101 & AGE <= 110 ~ "age101_110",
    AGE >= 111 & AGE <= 120 ~ "age111_120"
  )) %>% 
  group_by(age_range) %>% 
  #filter_all(any_vars(is.na(.))) 
  summarize(total = n()) 
  


age_totals
#write.csv(age_totals, file="age_totals.csv")
```

# birthplace totals 
```{r}

birthplace_totals <- df5 %>% 
  group_by(bpldname) %>%
  summarize(total = n()) 

birthplace_totals
#write.csv(birthplace_totals, file="birthplace_totals.csv")
```



