---
title: "merge_preprocess_master"
author: "Joseph Pacifico"
date: "2/9/2021"
output: html_document
---



```{r}
# bring in master, merge preprocess file

master <- openxlsx::read.xlsx("D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\data\\current_master_2_12_21.xlsx", sheet = 1)

master
```

master has 56,905 rows and 87 columns 



preprocess has 2614 rows and 79 columns 
```{r}
# bring in individual preprocess files exported to csv
preprocess <- read.csv('D:\\Libraries 1\\Documents\\QMSS\\Data For Good\\DataPreprocessingtoMaster\\preprocess2final2_16_2021.csv')

preprocess$X <- NULL
preprocess
```

```{r}
library(tidyverse)

as_tibble(preprocess$Notes.Other.Information)
```

# old master kick out variables 
```{r}

mastertest <- master 
# old master kick out variables 
mastertest <- mastertest %>% select(!c(Date.Mustered.Out,
                        Age.Months,
                        Directsur,
                        Directfirst,
                        Directocc,
                        Whitef,
                        Chinesem,
                        Chinesef,
                        Nativem,
                        Nativef,
                        Male.birth,
                        Female.birth,
                        Year.of.this.Record
                        ))
mastertest <- mastertest %>% 
  plyr::rename(c("Rank.(Military)" = "Rank_Military", 
                 "STREET" = "Street.Address",
                 "Employment" = "Employment.History",
                 "Relation.to.Head.of.Household" = "Relation.to.Head.of.House"))

mastertest
```
- master 74 columns 56905 before leftjoin

```{r}

preprocesstest <- preprocess
# preprocess kick out variables 
preprocesstest <- preprocesstest %>% select(!c(
                        Maiden.Name,
                        Marriage.Age,
                        Residence.Year,
                        Marital.status,
                        Cannot.Read..Write,
                        Employment.History,
                        Home.in.1861,
                        Able.to.Read.and.Write,
                        Middle.Name,
                        Dwelling.Number,
                        Able.to.Speak.English,
                        Certificate,
                        Number.of.Dwelling.in.Order.of.Visitation))

preprocesstest <- preprocesstest %>%
  plyr::rename(c("Surname" = "Last.Name",
                 "Given.Name" = "First.Name",
                 "Gender" = "Sex",
                 "Source" = "source",
                 "Birth.Place" = "PLACEOFBIRTH",
                 "Birth.Year" = "CalculatedBirthYear",
                 "Occupation" = "PROFESSION",
                 "Death.Date" = "Date.of.Death",
                 "DATE.OF.ENLISTMENT" = "Enlistment.Date",
                 "PLACE.OF.ENLISTMENT" = "Enlistment.Place",
                 "CAUSE.OF.DEATH" = "Cause.of.Death",
                 "Rank" =  "Rank_Military",
                 "Religion" = "RELIGION",
                 #"Street.Address" = "STREET",
                 #"Industry" = "PROFESSION",
                 #"Relationship" = "Relation.to.Head.of.House",
                 "State.Province" = "State/Province", 
                 "Marital.Status" = "MARITAL",
                 "Line.Number" = "Line", 
                 "Page.number" = "Page"
                 ))




preprocesstest$Joint.ID.for.Matched.Records <- as.character(preprocesstest$Joint.ID.for.Matched.Records)

preprocesstest$CalculatedBirthYear <- as.double(preprocesstest$CalculatedBirthYear)

preprocesstest$Age <- as.character(preprocesstest$Age)

mastertest$Date.of.Death <- as.character(mastertest$Date.of.Death)
mastertest$Cause.of.Death <- as.character(mastertest$Cause.of.Death)
mastertest$Rank_Military <- as.character(mastertest$Rank_Military)
mastertest$Enlistment.Date <- as.character(mastertest$Enlistment.Date)
mastertest$Enlistment.Place <- as.character(mastertest$Enlistment.Place)

masternew <- full_join(mastertest, preprocesstest)
masternew



```
- 109 columns and 59519 rows post left_join

master has 56,905 rows and 74 columns 
new master 59,519 rows, 109 columns 
35 columns added 

#check final_join for duplicate columns 
```{r}
# 35 columns not full_joined, double-check
masterlast <- masternew[,(ncol(masternew)-34):ncol(masternew)] # refers to the 4 last columns
masterlast
```
```{r}
unique(mastertest$Relation.to.Head.of.House)
unique(masterlast$Employment)
```

Relationship == Relation.to.Head.of.House x
Industry == PROFESSION x
Street == STREET x
NOTES == Notes.Other.Information 
Other.Information == Notes.Other.Information

Naturalization.Year > to notes x
Child > to notes x










```{r}
# unite Relationship == Relation.to.Head.of.House columns  
masternew <- masternew %>% 
  unite(Relation.to.Head.of.House,
        Relationship,
        Relation.to.Head.of.House,
        sep = " ") %>% 
  mutate(Relation.to.Head.of.House = gsub('NA', "", Relation.to.Head.of.House))

select(masternew, Relation.to.Head.of.House)
#unique(masternew$Relation.to.Head.of.House)
```



```{r}
# unite Industry == PROFESSION columns 
masternew <- masternew %>%
  unite(PROFESSION,
        PROFESSION,
        Industry,
        sep = " ") %>% 
  mutate(PROFESSION = gsub('NA', "", PROFESSION))
select(masternew, PROFESSION)
#unique(masternew$new_var)
```

```{r}
# unite Street == STREET columns 
masternew <- masternew %>%
  unite(Street.Address,
        Street.Address,
        Street,
        sep = " ") %>% 
  mutate(Street.Address = gsub('NA', "", Street.Address))
select(masternew, Street.Address)
#unique(masternew$new_var)
```


NOTES == Notes.Other.Information 
Other.Information == Notes.Other.Information
```{r}
# unite NOTES & Other.Information columns 
masternew <- masternew %>%
  unite(Notes.Other.Information,
        Notes.Other.Information,
        NOTES,
        Notes,
        Other.Information,
        sep = "|") %>% 
  mutate(Notes.Other.Information = gsub('NA', "", Notes.Other.Information))


masternew %>% select(Notes.Other.Information) %>% slice(57550)
masternew %>% select(Notes.Other.Information) %>% slice(59508)

#unique(masternew$new_var)
```

[1]   NA 1909 1864 1906 1900 1858 1898 1897 1870


```{r}
View(select(masternew, Notes.Other.Information))
```


# checking column values in master to see which are mostly empty 
```{r}

# to see which columns have an actual value or is just NAs
masterlast %>% summarise_all(funs(sum(!is.na(.))))

#https://stackoverflow.com/questions/50357668/count-number-of-non-na-values-for-every-column-in-a-dataframe
```

# columns to keep 
Maiden.Name 3
Father.s.Name 8
Mother.s.Name 8
Spouse 28
Child 8
City 35
MiddleName 46
Death.Place 71
Industry 8
Attended.School 29 
House.Number 25
Father.s.Birth.Place 73
Mother.s.Birth.Place 76
Street 35
Relationship 53
Immigration.Year 12
Father.of.Foreign.Birth 17
Mother.of.Foreign.Birth 17
Employment 2
Able.to.read 29
Able.to.Write 29
Family.Number 98
Residence.Place 12

To delete: 
Marriage.Age 2 x
Residence.Year 10 x
Marital.status 1 x
Dwelling.Number 74 x
Cannot.Read..Write 2 x
Able.to.Speak.English 32 x
Certificate 31 x
Number.of.Dwelling.in.Order.of.Visitation 18 x
Home.in.1861 7 x
Able.to.Read.and.Write 2 x

# export the new master
```{r}


write.csv(x=masternew, file="current_master_02_17_21.csv")
```


135 columns 59521 rows BEFORE
103 columns 59519 rows AFTER

- add in current master 2-12.xlsx
- check the last few added columsn 

columns that are empty in master:
Date.Mustered.Out 0 
Age.Months 9 
Directsur 0 
Directfirst 0 
Directocc 0 
Whitef 10 
Chinesem 0 
Chinesef 0 
Nativem 1
Nativef 1 
Male.birth 2
Female.birth 3 
Year.of.this.Record 0 
Maiden.Name 3 
Marriage.Age 2
Residence.Year 10 
Marital.status 1 
Cannot.Read..Write 2
Employment.History 2
Home.in.1861 7 
Able.to.Read.and.Write 2
Middle.Name 1



```{r}
colnames(master)
#unique(preprocess$MiddleName)
```




```{r}
notestest <- preprocess 
notestest <- notestest %>% unite(Notes.Other.Information1, 	
Naturalization.Year, sep = " ") 

notestest <- as.data.frame(paste(notestest$Notes.Other.Information, notestest$Notes.Other.Information1)) 

notestest <- notestest %>% plyr::rename(c('paste(notestest$Notes.Other.Information, notestest$Notes.Other.Information1)' = 'new_var'))

filter(notestest, "1863" %in% new_var)

notestest[grepl('1863', notestest$new_var), ] # check to see naturalization year == 1863 appears in the new notes column 






  mutate(new_var = gsub('NA', "", new_var)) %>% 
  mutate(new_var = gsub('_', "", new_var))%>% 
  #unite(Notes.Other.Information, NOTES, Other.Information)  %>%
  mutate(Notes.Other.Information = gsub('NA', "", Notes.Other.Information)) %>% 
  mutate(Notes.Other.Information = gsub('_', "", Notes.Other.Information)) %>% 
  unite(Notes.Other.Information, Notes.Other.Information, new_var, sep = " ")

masternew %>% select(Notes.Other.Information)
unique(preprocess$Naturalization.Year)
```

```{r}

colnames(master)

```